### 2021-04-29 paper.tex
sync with published version.

%% This is file `elsarticle-template-1-num.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public %
%License, either version 1.2 of this license or (at your option) any % later
%version.  The latest version of this license is in %
%http://www.latex-project.org/lppl.txt % and version 1.2 or later is part of all
%distributions of LaTeX % version 1999/12/01 or later.
%%
%% Template article for Elsevier's document class `elsarticle' % with numbered
%style bibliographic references
%%
%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $ % $URL:
%http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
%%
\documentclass[preprint,5p, twocolumn, authoryear]{elsarticle}

%% Use the option review to obtain double line spacing %
%\documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn % for a
%journal layout: % \documentclass[final,1p,times]{elsarticle} %
%\documentclass[final,1p,times,twocolumn]{elsarticle} %
%\documentclass[final,3p,times]{elsarticle} %
%\documentclass[final,3p,times,twocolumn]{elsarticle} %
%\documentclass[final,5p,times]{elsarticle} %
%\documentclass[final,5p,times,twocolumn]{elsarticle}

%% The graphicx package provides the includegraphics command.
\usepackage{graphicx}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amsmath}   % Extra math commands and environments from the AMS
\usepackage{amssymb}   % Special symbols from the AMS
\usepackage{amsthm}    % Enhanced theorem and proof environments from the AMS
\usepackage{latexsym}  % A few extra LaTeX symbols
\usepackage{array}
\usepackage{makecell}
\usepackage{units}
\usepackage{xcolor}
\usepackage[hidelinks=true]{hyperref}
\usepackage{natbib}

%\biboptions{authoryear,round}

%% The lineno packages adds line numbers. Start line numbering with %
%\begin{linenumbers}, end it with \end{linenumbers}. Or switch it on % for the
%whole article with \linenumbers after \end{frontmatter}.
\usepackage{lineno}
\usepackage[singlelinecheck=false]{caption}

% Define shortcuts in MPCSymbols.
\newcommand{\bbI}{\mathbb{I}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\norm}[1]{\left| #1 \right|}
\newcommand{\xhat}{\hat{x}}
\newcommand{\dhat}{\hat{d}}
\newcommand{\xseq}{\mathbf{x}}
\newcommand{\bbX}{\mathbb{X}}
\newcommand{\bbW}{\mathbb{W}}

%\biboptions{longnamesfirst, semicolon} % natbib.sty is loaded by default.
%However, natbib options can be % provided with \biboptions{...} command.
%Following options are % valid:

%%   round  -  round parentheses are used (default) %   square -  square
%brackets are used   [option] %   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option> %   semicolon  -  multiple
%citations separated by semi-colon %   colon  - same as semicolon, an earlier
%confusion %   comma  -  separated by comma %   numbers-  selects numerical
%citations %   super  -  numerical citations as superscripts %   sort   -  sorts
%multiple citations according to order in ref. list %   sort&compress   -  like
%sort, but also compresses numerical citations %   compress - compresses without
%sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

\newcommand{\underl}[1]{\underline{#1}} \newcommand{\overl}[1]{\overline{#1}}
\newcommand{\satlb}[1]{\textnormal{sat}(#1, \underline{#1}, \overline{#1})}
\newcommand{\tif}{\textnormal{if}}

\graphicspath{{./}{./figures/}{./figures/paper/}}

\newdefinition{assumption}{Assumption}
\newdefinition{prop}{Proposition}
\newdefinition{definition}{Definition}
\newdefinition{rmk}{Remark}
\newtheorem{thm}{Theorem}

\journal{Elsevier}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

\title{Industrial, large-scale model predictive control with structured neural
networks}

%% use the tnoteref command within \title for footnotes; % use the tnotetext
%command for the associated footnote; % use the fnref command within \author or
%\address for footnotes; % use the fntext command for the associated footnote; %
%use the corref command within \author for corresponding author footnotes; % use
%the cortext command for the associated footnote; % use the ead command for the
%email address, % and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}} % \tnotetext[label1]{}
\author[label1]{Pratyush Kumar\corref{cor1}}
\ead{pratyushkumar@ucsb.edu}
%% \ead[url]{home page} % \fntext[label2]{}
\cortext[cor1]{Corresponding Author}
%% \fntext[label3]{}

\author[label1]{James B. Rawlings}
\ead{jbraw@ucsb.edu}

\author[label2]{Stephen J. Wright}
\ead{swright@cs.wisc.edu}

\address[label1]{Department of Chemical Engineering, University of California, Santa Barbara, CA 93106, United States}
\address[label2]{Computer Sciences Department, University of Wisconsin-Madison, Madison, WI 53706, United States}

\begin{abstract}
The design of neural networks (NNs) is presented for treating large, linear
model predictive control (MPC) applications that are out of reach with available
quadratic programming (QP) solvers. First, we introduce a new feedforward
network architecture that enables practitioners to obtain offset-free
closed-loop performance with NNs. Second, we discuss the data generation
procedure to sample the state space relevant to training the NNs based on
anticipated online setpoint changes and plant disturbances. Third, we use the
input-to-state stability results available in the MPC literature and establish
robustness properties of NN controllers. Finally, we present illustrative
simulation studies on process control examples. We apply the NN design approach
and compare the performance with online QP based MPC on an industrial crude
distillation unit model with 252 states, 32 control inputs, and a control-sample
horizon length of 140. Parallel computing is used for data generation and
graphical processing units are used for network training. Anticipated plant
operational scenarios with setpoints and disturbances that may change during
operation must be sampled for NN training. After the offline design phase, NNs
execute MPC three to five orders of magnitude faster than an available QP solver
with less than $1\%$ loss in the closed-loop performance.
\end{abstract}

\begin{keyword}
Model predictive control \sep machine learning \sep deep neural networks \sep
large-scale systems \sep quadratic programming \sep inherent robustness
\end{keyword}

\end{frontmatter}

%% main text
\section{Introduction} \label{sec:introduction}

Model predictive control (MPC) is an online optimization based feedback control
technology. A dynamic model of the plant is used to make forecasts of the plant
measurements in response to the actuator movements, and an optimization problem
is solved in real time to determine the optimal actuator move to be applied to
the plant. For linear plant models, the optimization problem is a quadratic
program (QP). The development of efficient QP algorithms
\citep{kouzoupis:frison:zanelli:diehl:2018, wright:2019} allowed practitioners
to apply MPC in the process industries \citep*{qin:badgwell:2003, lahiri:2017}.

An alternative strategy to real time optimization for the deployment of MPC is
to characterize the MPC feedback law as a piecewise affine function defined on
polyhedral partitions of the state space, and use the feedback law online
\citep{bemporad:morari:dua:pistikopoulos:2002, seron:goodwin:dedona:2003}. The
real time computation in this approach is reduced to a table look-up in the
state space based on the current value of parameters in the QP such as the state
estimate and some steady-state targets that are computed for offset-free
control. The number of polyhedral partitions of the state space grows
exponentially with model dimensions, however, which prohibits the deployment of
this strategy on large-scale systems.

The MPC feedback law can be approximated using parametric functions such as
polynomials \citep{kvasnica:lofberg:fikar:2011}, various types of piecewise
affine functions \citep{bemporad:oliveri:poggi:storace:2011,
wen:ma:ydstie:2009}, and neural networks (NNs)
\citep{cavagnari:magni:scattolini:1999}. Among the function approximators, NNs
that use the rectified linear unit (ReLU) as the activation function have gained
attention recently \citep{chen:saulnier:atanasov:lee:kumar:pappas:morari:2018,
karg:lucia:2020, paulson:mesbah:2020, lovelett:dietrich:lee:kevrekidis:2020} to
approximate the MPC feedback law due to their ability to represent complex
piecewise affine functions \citep{montufar:pascanu:cho:bengio:2014} and execute
MPC faster in real time than QP solvers. The proposed NN design approach in
these works is to generate training data for the NN by solving QPs offline for a
set of feasible states of the QP, use the collected data to train a standard
feedforward network, then use the trained NN as the feedback controller online.

To make progress in the size of control applications achievable with MPC, large
problems in which QP solvers fail to deliver the control in the available sample
time should be addressed. The scalability of the NN approach to such problems
with large model dimensions and control horizon length in the QP has not been
demonstrated to date. Two issues arise in these problems: i) In a large
dimensional state space, the entire set of feasible states of the QP cannot be
sampled densely for the NN training. This issue is a form of the well-known
curse of dimensionality. ii) One may sample the state space partially to avoid
this problem, but the time required for the data generation can still be
impractical due to the long time required to solve a QP for each sampled state.
We demonstrate in this article by case studies that the issue (ii) can be
resolved for many size of problems with parallel computing, and the issue (i) by
sampling the state space for typical plant operational scenarios. The following
feature must be present in the large-scale application of interest for the
partial sampling scheme proposed in this article to be applicable: For a
particular operating mode of the plant, the number of frequently changing
setpoints and large magnitude disturbances must be small enough such that only a
small fraction of the entire large dimensional state space is visited during
that operation mode. The MPC feedback law can then be reliably approximated
using NNs for the different modes of plant operations, each driven by their
respective set of setpoint changes and disturbances.

From the above viewpoint, this article discusses the design of NNs as an
alternative to real time optimization in MPC, with an emphasis on large
applications that are challenging for QP solvers. We focus on setpoint tracking
MPC problems relevant to the process industries. We present a new feedforward
network architecture to obtain offset-free closed-loop performance with the NNs.
Additionally, we use the input-to-state stability results available in the MPC
literature, and establish conditions under which NN controllers are robust to
state estimation errors and process disturbances. We present case studies on
process control examples that demonstrate the scalability of the proposed NN
design approach.

The related work of \cite{chen:wang:atanasov:kumar:morari:2019} demonstrates the
scalability of NNs to approximate the MPC feedback law on systems with state
dimensions up to 36, on the control problem of regulation to the origin. By
contrast, our paper considers setpoint tracking offset-free MPC problem and
state dimensions up to 250. \cite{drgona:picard:kvasnica:helsen:2018} take a
different approach for the scalability of NN controllers, in which the
dimensionality of the parameters in the MPC QP is first reduced using principal
component analysis, and the NN controller is developed in a low dimensional
state space.

The rest of this article is organized as follows. In the next section, we
briefly review offset-free linear MPC. In Section \ref{sec:controller_design},
we discuss the NN controller design procedure. We present the robustness
property of NNs in Section \ref{sec:robustness}. Section
\ref{sec:simulation_studies} presents simulation studies on two large process
control examples and some concluding thoughts appear in Section
\ref{sec:conclusion}.

Notation. The symbols $\bbI$ and $\bbR$ are used to denote integers and reals
respectively. Subscripts denote restrictions (e.g., $\bbR_{\geq 0}$ for
nonnegative reals and $\bbI_{a:b}$ for integers in the closed interval $[a,
b]$). The Euclidean norm is denoted by $\norm{\cdot}$. A function $\alpha :
\bbR_{\geq 0} \rightarrow \bbR_{\geq 0}$ is of class $\mathcal{K}$ if it is
continuous, zero at zero, and strictly increasing. This function is of class
$\mathcal{K}_{\infty}$ if it is of class $\mathcal{K}$ and unbounded ($\alpha(s)
\rightarrow \infty$ as $s \rightarrow \infty$). A function $\beta : \bbR_{\geq
0} \times \bbI_{\geq 0} \rightarrow \bbR_{\geq 0}$ is of class $\mathcal{KL}$ if
it is continuous, for each $k \in \bbI_{\geq 0}$, $\beta(\cdot, k)$ is of class
$\mathcal{K}$, and for each $s \geq 0$, $\beta(s, \cdot)$ is nonincreasing and
satisfies $\textnormal{lim}_{k \rightarrow \infty} \beta(s, k) = 0$. Given $V :
X \rightarrow \bbR_{\geq 0}$ and $\tau > 0$, define $\textnormal{lev}_{\tau} V =
\{ x \in X \mid V(x) \leq \tau \}$. Bold symbols, e.g., $\mathbf{d}$ denote
sequences, $d(k)$ denotes an element of $\mathbf{d}$ at time $k \in \bbI_{\geq
0}$, and $\mathbf{d}_i$ denotes the collection of elements of $\mathbf{d}$ for
$k \in \bbI_{0:i-1}$. Define $\norm{\norm{\mathbf{d}_i}} = \textnormal{max}_{k
\in \bbI_{0:i-1}} \norm{d(k)}$. For a given vector $a$, lower and upper bounds
($\underl{a}, \overl{a}$), define the saturation function as $\satlb{a} = \{ a \
\tif \ \underl{a} \leq a \leq \overl{a}; \underl{a} \ \tif \ a < \underl{a};
\overl{a} \ \tif \ \overl{a} < a \}$.

\section{Linear model predictive control} \label{sec:mpc}

\subsection{Model}
We consider a linear discrete time model augmented with an integrating
disturbance model \citep{pannocchia:rawlings:2003}:
\begin{align}\label{eq:ltimodel}
    x^+ = Ax + Bu + B_dd, \quad d^+ = d, \quad y = Cx + C_dd,
\end{align}
in which $x \in \bbR^n$ is the state, $u \in \bbR^m$ is the control input, and
$d \in \bbR^d$ is the state of the disturbance model. The matrices $A \in
\bbR^{n \times n}$, $B \in \bbR^{n \times m}$,  
$C \in \bbR^{p \times n}$ denote the actuator to measurement model; and $B_d \in
\bbR^{n \times d}$, $C_d \in \bbR^{p \times d}$ is the disturbance model. The
objective of the disturbance model is to remove offset in the controlled plant
measurements at steady state operations and maintain the controlled variables at
their setpoints. We assume in the rest of this article that a Kalman filter can
be constructed to estimate the model state ($\xhat$) and disturbance ($\dhat$)
from measurements.

\subsection{Target selector}
Given the disturbance estimate, the input and controlled measurement setpoints,
we consider the following target selector QP:
\begin{align}
    \min_{x_s, \ u_s}  \dfrac{1}{2} |u_{sp} - u_s|^2_{R_s}
\end{align}
subject to
\begin{align}  
    \begin{bmatrix}
        I -A & -B \\
        HC & 0
    \end{bmatrix} \begin{bmatrix}
        x_s \\
        u_s
    \end{bmatrix} &= \begin{bmatrix}
        B_d\dhat \\
        r_{sp} - HC_d\dhat
    \end{bmatrix}, \\
    \underline{u} \leq u_s \leq \overline{u},
\end{align}
in which ($x_s, u_s$) is the target steady-state pair, $r=Hy$ are the controlled
measurements chosen as some subset or a linear combination of all the
measurements, ($\underline{u}, \overline{u}$) are the actuator bounds, and
($u_{sp}, r_{sp}$) are the input and controlled measurement setpoints. The
equality constraint $r_{sp} = H(Cx_s + C_d\dhat)$ can be difficult to satisfy
exactly in real time depending on the controlled measurement setpoint and
disturbance estimate. In this case, the hard setpoint equality constraint can be
relaxed and moved to the stage cost, such that the target selector computes a
steady state to minimize the offset in the controlled measurements. We assume
for the simulation studies in this article that the input setpoint ($u_{sp}$) is
fixed at some chosen steady state and only the controlled measurement setpoint
($r_{sp}$) changes in real time.

\subsection{Regulator}
Based on the state estimate and target steady-state pair, the following QP is
solved to determine the actuator move to be applied to the plant:
\begin{align} \label{eqn:regulator}
    V_N(\tilde{x}(0), \mathbf{\tilde{u}}) = \dfrac{1}{2}\sum_{k=0}^{N-1} & 
    \Big(|\tilde{x}(k)|^2_Q  + |\tilde{u}(k)|^2_R \Big) + 
    \dfrac{1}{2}|\tilde{x}(N)|^2_P
\end{align}
subject to 
\begin{align}
    \tilde{x}^+ &= A\tilde{x} + B\tilde{u}, \quad \tilde{x}(0) = \xhat - x_s, \label{eqn:model_eq} \\
    &\underline{u} \leq \tilde{u} + u_s \leq \overline{u},
    \label{eqn:regulator_constraints}
\end{align}
in which $\xhat$ is the state estimate after the current measurement, $N$ is the
control horizon length, and ($\tilde{x}, \tilde{u}$) are the state and control
in deviation from the current steady-state targets ($x_s, u_s$). The penalty
matrix $P$ is chosen as the optimal cost-to-go matrix of the unconstrained
infinite horizon linear quadratic problem. The decision variable in this QP is
the control sequence $\mathbf{\tilde{u}} = [\tilde{u}(0)', \tilde{u}(1)', ...,
\tilde{u}(N-1)']'$, and the state sequence $\tilde{\xseq}$ is fully determined
by the model equality constraints \eqref{eqn:model_eq}. The first move of the
optimal solution is denoted as the MPC feedback law, $\kappa_N(\xhat, x_s, u_s)
= \mathbf{\tilde{u}}^0(0;\xhat, x_s, u_s) + u_s$, which is applied to the plant.
This feedback law is a function of the state estimate and target steady-state
pair. A rate-of-change penalty on the control can be imposed in the QP objective
by augmenting the dynamic model with a surrogate state for the control as
discussed in \cite{rao:rawlings:1999}. In this case, the control injected at the
previous timestep ($u_{-1}$) is supplied to the QP, and the feedback law becomes
a function of the previous control, in addition to the state estimate and target
steady-state pair.

Several algorithms have been proposed in the literature over the years to solve
the above MPC regulator QP. \cite{kouzoupis:frison:zanelli:diehl:2018} and
\cite{wright:2019} give recent reviews on the development of convex optimization
algorithms to solve this linear MPC QP. For the simulation studies in this
article, we eliminate the state trajectory from the set of decision variables
and formulate a dense QP. The decision variable in the dense QP is the future
control trajectory $\mathbf{\tilde{u}}$, and the QP solver CVXOPT
\citep{vandenberghe:2010}, which is tailored for dense problems is used to
generate data for NN training and for timing comparisons with online QP based
MPC. Any improvement in the QP formulation and solution algorithm is
advantageous for both the online QP based MPC and NN approaches, as fast QP
solvers reduce the offline data generation time required for the design of NNs.

\section{Neural network design} \label{sec:controller_design}

In the offset-free linear MPC algorithm discussed in the previous section, the
target selector QP is usually small compared to the regulator QP, which consumes
most of the online computation time. Therefore we focus on designing a NN that
approximates the MPC feedback law ($\kappa_N(x, x_s, u_s)$) for an operationally
relevant set of states and steady-state targets, such that the NN can be used
online as the feedback controller for the plant in place of solving the MPC
regulator QP.

\subsection{Structured neural network}
An intuitive approach is to build a feedforward NN that takes the triple ($x,
x_s, u_s$) as its input and outputs a control that is close to the optimal
control. This strategy has been proposed in \cite{karg:lucia:2020} and 
\cite{chen:saulnier:atanasov:lee:kumar:pappas:morari:2018}
for a single fixed steady state. The approach has not been demonstrated to scale
to the size of model dimensions and the setpoint tracking offset-free MPC
problem considered in this article. An issue with this strategy is that the NN
has no knowledge about the MPC feedback law, and can require large amounts of
training data in large state dimensions to obtain a reasonable controller. At a
minimum, the MPC feedback law is such that at the origin, the control action
taken by the MPC controller is zero. For the setpoint tracking MPC problem
considered in this article, the NN control law must be such that at steady-state
operations, the NN maintains the plant at the desired steady state such that the
controlled measurements are at their setpoints. This structure implies $u_s =
\kappa_{NN}(x = x_s, x_s, u_s)$, in which we use $\kappa_{NN}(\cdot)$ to denote
the NN control law. To incorporate this information in a NN, we introduce the
following architecture:
\begin{align}
z_0 &= \begin{bmatrix}
    x', & x_s', & u_s', & x_s', & x_s', & u_s' \end{bmatrix}', \notag \\
    f_i(z_{i-1}) &= \begin{bmatrix}
        W_i & 0 \notag \\
        0 & W_i
          \end{bmatrix}z_{i-1}  + \begin{bmatrix}
            b_i \\
            b_i
            \end{bmatrix}, \notag\\
    z_i &= g(f_i(z_{i-1})), \textnormal{for} \ i \in \bbI_{1:h}, \notag \\
    u &= u_s + \begin{bmatrix}
  W_{h + 1}, & -W_{h + 1} \\
    \end{bmatrix} z_h,
\label{eq:NNstruct}
\end{align}
in which $g(a) = \textnormal{max} (0, a)$, is the ReLU operation applied
element-wise on its vector input, $z_i$ is the output of each hidden layer of
the network, $z_0$ is the input to the network, $h$ is the number of hidden
layers, and $i \in \bbI_{1:h}$. The parameters to be optimized using training
data are ($W_i, b_i$) and $W_{h+1}$, which are matrices of appropriate
dimensions. At steady state, when $x=x_s$ in $z_0$, \eqref{eq:NNstruct} outputs
$u_s$ regardless of the choice of weights and the activation function, as the
output of all hidden layers ($z_i$) have same subvector repeated in the upper
and lower half and the term $\begin{bmatrix} W_{h + 1}, & -W_{h+ 1} \\
      \end{bmatrix} z_h$ equals zero. The NN structure \eqref{eq:NNstruct} is
equivalent to the difference in the outputs of a standard feedforward NN with
inputs ($x, x_s, u_s$) and ($x_s, x_s, u_s$) respectively, with $u_s$ added to
the final output. The software tensorflow
\citep{abadi:agarwal:barham:brevdo:et-al:2015} is used for NN training; the
structured architecture can be built with symbolic differentiation tools
available in tensorflow.

When a rate-of-change penalty on the control is included in the regulator QP,
the input to the structured NN is modified to:
\begin{align*}
    z_0 &= 
    \begin{bmatrix}
    x', & u_{-1}', & x_s', & u_s', & x_s', & u_s', & x_s', & u_s' \\
    \end{bmatrix}',
\end{align*}
in which $u_{-1}$ is the previous control applied to the plant and the
dimensions of the matrices ($W_i, b_i$) and $W_{h+1}$ are appropriately
adjusted. Even after training, the control produced by the NN may not satisfy
the hard actuator constraints, so the saturation function
($\textnormal{sat}(\cdot, \underline{u}, \overline{u})$) is applied to the
output of the NN.

\subsection{Data generation}

For one fixed steady state, the set of feasible states for the MPC QP defined by
\eqref{eqn:regulator}-\eqref{eqn:regulator_constraints} is the entire
$n$-dimensional state space $\bbR^n$. Additionally, the target steady state may
change during plant operation, depending on the setpoint and disturbance
estimate. The domain of the MPC control law ($\kappa_N(x, x_s, u_s)$) is
therefore the state space $\bbR^{n}$ and the possible set of steady-state target
pairs. Sampling this entire domain of the MPC control law for NN training is
difficult in large dimensions; we show in Section \ref{sec:simulation_studies}
that it is not required for practical applications.

A majority of large-scale petrochemical plants operate in a relatively small
number of operating regimes or scenarios. Each operational scenario is driven by
a selected few controlled measurement setpoints that depend on product demands
and some large magnitude disturbances that may change frequently, while the
setpoints for other measurements remain constant for long periods of time. The
states and steady-state targets spanned in a closed-loop operation per scenario
is thereby considerably less than the entire domain of the MPC control law.
These characteristics of typical operation of large chemical plants can be
exploited, and we propose the following data generation procedure to perform
offline simulations using the model to collect the operationally relevant set of
($x, x_s, u_s, \kappa_N(\cdot)$) for NN training.

\begin{enumerate}
    \item Determine the anticipated range of controlled measurement setpoint
    changes ($r_{sp}$) to be made during the plant operational scenario. We
    denote this range of setpoints as ($\underline{r}_{sp}$,
    $\overline{r}_{sp}$).
    \item Identify the set of physical disturbances ($d$) that may affect the
    plant and their range of values ($\underline{d}$, $\overline{d}$). We use
    physical disturbances in this article and assume that a disturbance model
    identification is performed to obtain the matrices $B_d$ and $C_d$.
    \item Create pseudo random binary signals (PRBS) of $r_{sp}$ and $d$ in
    their respective range of values.
    \item Initialize the model at some chosen steady state. Perform an offline
    simulation with the generated PRBS signals by solving the target selector
    and regulator QPs for the transient states and steady-state targets
    encountered in the simulation. 
\end{enumerate}

During plant operation, it is expected that the NN will encounter states and
steady-state target pairs not used in training. But similar values are expected,
since an adequate model and appropriate range of anticipated setpoints and plant
disturbances are used for the data generation. During the closed-loop operation,
the NN produces an interpolated actuator move that is applied to the plant.

As we consider large MPC problem sizes in this article for the NN controller,
the data generation step can take impractical amounts of time if performed
serially, due to the long time (minutes) required to solve each QP. For these
large problems, several PRBS signals of the setpoints and disturbances are
generated and multiple simulations are performed in parallel over several cores
in a CPU and across multiple CPUs as well. The data generated from all the
simulations are collected and used for the NN training.

\subsection{Training}

The following mean squared error (MSE) is minimized to determine the weights for
the NN controller:
\begin{align}
    J (\theta) = \dfrac{1}{2}\sum_{j=1}^{N_{tr}} 
[\kappa_{NN}(x_j, x_{sj}, u_{sj}; \theta) - \kappa_N(x_j, x_{sj}, u_{sj})]^2 
\end{align}
in which $\theta$ is the set of parameters, ($W_i, b_i, W_{h+1}$) for $i \in
\bbI_{1:h}$, subscript $j$ is used to denote the training sample index, and
$N_{tr}$ is the number of training samples. The stochastic optimization
algorithm Adam \citep{kingma:ba:2014} is used for all the case studies in this
article. We do not consider any regularization penalty in the training
objective. The recent empirical and theoretical works in the machine learning
literature have shown that NNs with large number of parameters have good
interpolation capabilities \citep{belkin:hsu:ma:mandal:2019,
zhang:bengio:hardt:recht:vinyals:2017, arora:simon:hu:li:wang:2019,
allen-zhu:li:liang:2019} without any form of explicit regularization. The
generalization abilities of NNs for the architectures considered in this article
is examined in the simulations presented in the Section
\ref{sec:simulation_studies}.

\section{Robustness of neural networks} \label{sec:robustness}

\begin{figure*}[!h]
    \centering
	\includegraphics[page=1, width=0.7\textwidth,
		height=0.46\textheight]{cstrs_comparision_plots.pdf}    
        \caption{Transient closed-loop actuator trajectories of the CSTRs with
        flash separator example with the NN and optimal MPC controllers. The
        performance of the NN and optimal MPC controllers are almost
        indistinguishable.}   
    \label{fig:cl_cstrs_inputs}
\end{figure*}

\begin{figure*}[!h]
    \centering
	\includegraphics[page=2, width=0.7\textwidth,
		height=0.46\textheight]{cstrs_comparision_plots.pdf} \caption{Transient
		closed-loop controlled measurement trajectories of the CSTRs with flash
		separator example with the NN and optimal MPC controllers. The
		performance of the NN and optimal MPC controllers are almost
		indistinguishable.}
		\label{fig:cl_cstrs_outputs}
\end{figure*}

We now present the robustness properties of NN controllers. The optimal MPC
controller designed for a nominal linear system ignoring disturbances is known
to be inherently robust to bounded state estimation errors and disturbances
\citep{heath:wills:2005, pannocchia:rawlings:wright:2011}. The approximation
error of the MPC feedback law by a NN can be viewed as an additional
disturbance, and existing input-to-state stability results
\citep{sontag:wang:1995} can be used to establish the allowable approximation
error in the NN for the NN controller to be robust to disturbances. Theorem
\ref{thm:nnrobustness} below states this allowable approximation error and
specifies the size of disturbances for which the NN controller is robust. This
result also holds for other approximate MPC controllers, which use different
parametric functions to approximate the optimal control law, such as polynomials
and other piecewise affine functions.

The related work by \cite{hertneck:kohler:trimpe:allgower:2018} examines the
robustness of nonlinear systems in feedback with approximate MPC controllers.
The MPC control law approximation error is treated as an input disturbance, and
a robust MPC formulation based on constraint tightening procedures is used to
establish the allowable approximation error. The robustness of NN controllers
trained with samples of nominal MPC control laws in presence of process
disturbances and state estimation errors has not been established to date in the
literature.

The problem of steering the state of the linear system (\ref{eq:ltimodel}) to
one fixed steady state ($x_s, u_s$) is considered. For the analysis in this
section, the state and control are transformed in deviation variables as $x := x
- x_s$ and $u := u -u_s$, and we define the control problem as regulation to the
origin. The optimal MPC control law is denoted as $\kappa_N(x)$, and the NN
control law as $\kappa_{NN}(x) = \kappa_N(x) + e_{NN}(x)$, in which $e_{NN}(x)$
is the approximation error of the optimal control law by the network. Assume in
the analysis that the target steady state is such that the unconstrained linear
quadratic regulator (LQR) satisfies the actuator constraints in a small
neighborhood near the origin.

As reviewed in \cite{mayne:rawlings:rao:scokaert:2000}, the nominal stability of
MPC can be established using the optimal cost function ($V_N^0(x)$) of the QP as
a Lyapunov function. The region of attraction of the closed-loop system $x^+ =
Ax + B\kappa_N(x)$ when no hard terminal region constraint is included in the QP
can be characterized as $\mathcal{X}_N = \textnormal{lev}_{Nd + \tau} \
V_N^0(x)$ \citep{limon:alamo:salas:camacho:2006}. The parameter $\tau$ is the
level set parameter of the terminal region chosen as $\bbX_f =
\textnormal{lev}_{\tau} \ x'Px$, $d > 0$ is a constant such that $\ell(x, u)
\geq d$ for all $x \in \bbR^n \setminus \bbX_f$, and $\underline{u} \leq u+u_s
\leq \overline{u}$. The parameter $\tau$ is chosen small enough such that the
unconstrained LQR satisfies the actuator constraints for all $x \in \bbX_f$. The
optimal cost function is continuous and satisfies $c_1\norm{x}^2 \leq V_N^0(x)
\leq c_2\norm{x}^2$ and $V_N^0(x^+) - V_N^0(x) \leq -c_1\norm{x}^2$ for some
$c_2 \geq c_1 > 0$.

Assume that a state estimate ($\xhat$) is used by the NN to compute the control.
To analyze robustness, the following perturbed linear system is considered:
\begin{align} \label{eq:perturbed_sys}
    \xhat^+ = A\xhat + B\kappa_N(\xhat) + Be_{NN}(\xhat) + w - Ae + e^+,
\end{align}
in which $\xhat = x + e$ is the state estimate, $w$ is a process disturbance,
and $e$, $e^+$ are the state estimation errors at the current and next timestep,
respectively. We use $\phi_d(k;\xhat)$ to denote a solution of the perturbed
system  at time $k$ starting from an initial state $\xhat$; and $\phi(k;x)$ to
denote a solution  of the nominal system starting  from an initial state $x$
when the state estimation errors and process disturbances are zero.

\begin{thm} \label{thm:nnrobustness} For all $ 0 < \rho \leq Nd + \tau$, there
exist constants $\delta_1, \delta_2, \delta_3 > 0$, functions $\beta(\cdot) \in
\mathcal{K} \mathcal{L}$ and $\alpha_e(\cdot) , \alpha_w(\cdot),
\alpha_{n}(\cdot) \in \mathcal{K}$, such that for all disturbance sequences
satisfying $\norm{\norm{\mathbf{e}_{k+1}}} \leq \delta_1$,
$\norm{\norm{\mathbf{w}_k}} \leq \delta_2$, and $\norm{e_{NN}(\xhat)} \leq
\delta_3$, and for all $\xhat$ in the set $S := \textnormal{lev}_{\rho} \
V_N^0(\xhat)$, we have the bound $\norm{\phi_d(k; \xhat)} \leq
\beta(\norm{\xhat}, k) + \alpha_e(\norm{\norm{\mathbf{e}_{k+1}}}) +
\alpha_w(\norm{\norm{\mathbf{w}_k}}) + \alpha_n(\bar{e}_{NN})$, in which
$\bar{e}_{NN} = \textnormal{max}_{\xhat \in S} \norm{e_{NN}(\xhat)}$.
\end{thm}

The proof of this theorem and the required input-to-state stability definitions
are provided in Appendix \ref{app:theorem1}. In the nominal case, the
disturbance sequences $\mathbf{e}$ and $\mathbf{w}$ are zero, and the
closed-loop states have the bound $\norm{\phi(k; x)} \leq \beta(\norm{x}, k) +
\alpha_n(\bar{e}_{NN})$. An additional value of the NN training is to reduce the
worst approximation error ($\bar{e}_{NN}$) in the state space of interest, thus
reducing the bound on the closed-loop states.

\section{Simulation studies} \label{sec:simulation_studies}

In this section, two case studies are presented that demonstrate the scalability
of the proposed NN design approach. The closed-loop performance of NNs and
computational benefits over online QP based MPC are analyzed in these case
studies. After the offline design of NNs, two types of validation simulations
are performed: (i) with setpoints and disturbances not present in the training
data but within the same range of values used to generate the training data,
(ii) with extra setpoint signals assumed to be unknown during the offline
network design phase. Both these types of simulations shed light on the
interpolation and extrapolation abilities of the NNs, and are referred
subsequently as simulations with \textit{expected} and \textit{unexpected} plant
behaviors, respectively.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth,
                     height=0.25\textheight]{cstrs_with_flash.pdf}
    \caption{Schematic of the CSTRs with a flash separator model.}
    \label{fig:schematic_cstrs}
\end{figure}

The validation simulations are performed directly with the plant model for our
case studies. For industrial deployment, the NN training should be followed by
an offline validation step of the trained network to verify the quality of the
NN controller prior to online deployment. This validation can be performed by:
a) quantifying the MSE of the trained NN controller on test data generated using
the optimal MPC controller, b) examining closed-loop performance metrics in
simulations carried out with the linear model used in MPC and the optimal MPC
and NN controllers, similar to the simulations performed in this section with
the plant model. Additional probabilistic validation techniques for NNs trained
using robust MPC solutions for nonlinear models are proposed in
\cite{karg:alamo:lucia:2019} and \cite{hertneck:kohler:trimpe:allgower:2018}.

The closed-loop performance of the following alternative and equivalently fast
controllers as NNs are also compared in the validation simulations: (a) steady
state controller (SS): the solution of the target selector is directly applied
to the plant $u = u_s$; (b) saturated linear quadratic regulator (satK): the
unconstrained LQR gain $K$ is computed and used in the control law $u =
\textnormal{sat}(K(x-x_s) + u_s, \underline{u}, \overline{u})$; (c) short
horizon controller (SH): an MPC problem with a short control horizon is solved
online; (d) Unstructured NN (NN-UNS): a standard feedforward network is trained
with the triple ($x, x_s, u_s$) as the input and then used as the feedback
controller online.

To gauge the performance of NNs in the validation simulations, the offline
computational effort required to build the NNs, and the memory required for the
deployment of NNs, the following metrics are examined:

\begin{itemize}
    \item Controller performance index
    \begin{align*}
    \Lambda_k = \dfrac{1}{k}\sum_{t=1}^{k} & \big(\norm{x(t) - x_{s} (t)}^2_{Q} \\
        & + \norm{u(t) - u_{s} (t)}^2_{R} + \norm{\Delta u(t)}^2_{S}\big);
    \end{align*}
    \item \% Performance loss = $100(\Lambda^{\textnormal{F}}_{N_{t}} -
    \Lambda^{\textnormal{MPC}}_{N_{t}})/\Lambda^{\textnormal{MPC}}_{N_{t}}$;
    \item Average and worst-case speedups;
    \item Data generation and NN training times;
    \item Memory required to store the weights of the NNs.
\end{itemize}
Here, $N_t$ is the number of simulation time steps, and
$\Lambda^{\textnormal{F}}_{N_{t}}$ and $\Lambda^{\textnormal{MPC}}_{N_{t}}$ are
the average stage costs obtained at the end of the simulation period by the fast
controllers and the MPC controller. The data generation and online timing
comparisons are performed on a computing cluster that has several multi-core
CPUs of clock speed 2.4GHz. The NN trainings are performed on a Tesla V100-SXM2
GPU that has 32 GB memory. Code for the simulation studies is available at:
\url{https://github.com/pratyushkumar211/industrial_nnmpc_2021}.

\begin{table*}[t]
    \caption{Summary of the simulation study for the CSTRs with a flash
    separator example.}   
    \begin{tabular}{ |c|c|c|c|c|c|c|c| }
        \hline
        Controller & \thead{Neural network \\ architecture} & \thead{Number of
        \\ Parameters} & \thead{Memory \\ footprint \\ (MB)} & \thead{Training
        \\ time \\ (min)} & \thead{\% Performance\\ losses \\
        Expected, Unexpected \\ plant behaviors} & \thead{Average \\ speedup} &
        \thead{Worst case \\ speedup} \\
        \hline
    SS &  &  & &  & 85.18\%, 106.39 \% &  & \\ 
    satK &  &  &  &  & 41.03\%, 27.61 \% & $2.47 \times 10^5$ & $3.07 \times
    10^5$ \\ 
    SH &  &  &  &  & 1.61\%, 2.46 \% & $3.75 \times 10^3$ & $8.4 \times 10^1$ \\

    NN-UNS & [36, 224, 224, 224, 6] & $1.10 \times 10^5$ &  &  & 80.49\%, 73.29
    \% & $9.31 \times 10^4$ & $4.99 \times 10^3$ \\ 
NN-3-448 & [72, 448, 448, 448, 6] & $1.10 \times 10^5$ & 0.84 & 13.02 & 0.28 \%,
5.57 \% & $5.26 \times 10^4$ & $1.45 \times 10^3$ \\ 
NN-3-480 & [72, 480, 480, 480, 6]  & $1.26 \times 10^5$ & 0.96  & 12.51 & 0.34
\%, 7.58\% & $5.09 \times 10^4$ & $1.53 \times 10^4$ \\ 
NN-3-512 & [72, 512, 512, 512, 6]  & $1.42 \times 10^5$ & 1.08  & 12.77 & 0.16
\%, 5.90 \% & $4.62 \times 10^4$ & $1.58 \times 10^4$ \\ 
        \hline
       \end{tabular}
       \label{table:cstrs}      
\end{table*}

\begin{figure*}[!h]
    \includegraphics[page=31, width=0.5\textwidth,
        height=0.27\textheight]{cstrs_comparision_plots.pdf} \hfill
        \includegraphics[page=34, width=0.5\textwidth,
        height=0.27\textheight]{cstrs_comparision_plots.pdf} \vfill
    \vspace{-0.1in}
    \begin{center}\includegraphics[page=33, width=0.5\textwidth,
        height=0.27\textheight]{cstrs_comparision_plots.pdf}
    \end{center}
    \vspace{-0.2in}
    \caption{Simulation statistics for the  
    CSTRs with a flash separator example. Data requirement to obtain accurate NN
    controllers (top left), online computation times of the NNs and MPC
    controller (top right), and the controller performance index obtained with
    the NNs, SS, satK and MPC controllers (center bottom).The performance of the
    three NN and optimal MPC controllers are indistinguishable.}	
    \label{fig:cstrs_statistics}
\end{figure*}

\subsection{CSTRs with a flash separator}

For the first example, the plant consists of two stirred tank reactors in series
with a flash separator, depicted in Figure \ref{fig:schematic_cstrs}. In both
the CSTRs, a desired product $B$ is produced with the reaction $A \rightarrow
B$, and an undesired product $C$ is produced with the reaction $B \rightarrow
C$. The reactant $A$ is the major component in the feed streams supplied to the
reactors. After the second CSTR, the reaction mixture is sent to the
non-adiabatic flash that separates the reactant $A$ from the product $B$. The
$A$ rich vapor phase is recycled back to the first CSTR. The plant is simulated
using the set of nonlinear ordinary differential equations (ODEs) available in
\cite{venkat:2006}, Appendix. The model has 12 states ($H_r, x_{Ar}, x_{Br},
T_r, H_m, x_{Am}, x_{Bm}, T_m, H_b, x_{Ab}, x_{Bb}, T_b$), 6 control inputs
($F_0, F_1, D, Q_r, Q_m, Q_b$), and 5 disturbances ($x_{A0}, x_{A1}, x_{B0},
x_{B1}, T_0$). The controlled measurements with setpoints are the heights and
temperatures of the three units, and we assume that all the states are measured.
The sample time for the measurements is chosen as 10 seconds. The parameter
values used to simulate the ODEs for the plant, the actuator constraints, and
the bounds of the controlled measurements and disturbances used for the data
generation and validation simulations are shown in Table \ref{table:cstrs_pars},
Appendix \ref{app:cstrs_pars}.

A linear discrete time model with the chosen sample time is obtained at the
steady state shown in the Table \ref{table:cstrs_pars}. This model, inputs and
outputs are scaled for the MPC controller such that the input constraints
satisfy $\overline{u} - \underline{u} = \mathbf{2}$. The tuning parameters and
control horizon length for the regulator are chosen as $Q = 10^3 C'C$, $R = 0.1
I$, $S = 0.1 I$, and $N = 450$. A total $1.5 \times 10^5$ samples of ($x,
u_{-1}, x_s, u_s, \kappa_N(\cdot)$) is generated for NN training, parallelized
in 100 separate offline simulations. The time consumed for this data generation
step was 4.2 hours. The setpoint of the height of the second CSTR ($H_m$) is
assumed to be fixed during this data generation process, and is included as an
additional setpoint in the validation simulations for the unexpected plant
behavior case.

Three structured NN architectures are considered. Each NN has three hidden
layers with 448 (NN-3-448), 480 (NN-3-480), and 512 (NN-3-512) nodes in each
hidden layer respectively. For validation simulations, two sets of setpoint and
disturbance signals with 24 setpoint changes and 48 disturbance changes are
generated for 4320 timesteps (12 hours), one set each for the expected and
unexpected plant behavior cases. The short horizon MPC controller is constructed
with a horizon length of $N = 10$, and the architecture of the unstructured NN
consists of three hidden layers with 224 nodes in each hidden layer. An
additional scaling of $x$ and $x_s$ is performed for the NN training as $x :=
2x/(x_{\textnormal{max}} - x_{\textnormal{min}})$ and $x_s :=
2x_s/(x_{\textnormal{max}} - x_{\textnormal{min}})$, in which
$x_{\textnormal{max}}$ and $x_{\textnormal{min}}$ are the maximum and minimum
values of the state observed in the training data. This scaling is also
performed in real time when the NN is used to compute the actuator move.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.45\textwidth, height=0.25\textheight]{cdu.pdf}
    \caption{Schematic of the industrial crude distillation unit model.}
    \label{fig:schematic_cdu}
\end{figure}

First, we study the amount of data required to obtain an accurate NN controller
is studied. To this aim, NNs with the three chosen architectures are trained
with incremental increases of $10^4$ training samples, ranging from $4 \times
10^4$ to $1.5 \times 10^5$ samples. A batch size of 1024 samples is used in
Adam, and all the NNs are trained for 2000 epochs. For all the NN trainings
performed, 10$\%$ of the training dataset is kept as a buffer dataset, which is
monitored during training, and the network weights are updated only when the MSE
decreases on this buffer dataset after every training epoch. The NNs after
training are used in validation simulations with the setpoint and disturbance
signals generated for the expected plant behavior case. The closed-loop
performance is compared with the optimal MPC controller and the $\%$ performance
loss obtained is plotted in Figure \ref{fig:cstrs_statistics} (top left) as a
function of the number of training samples. All the trained NNs provide less
than $1 \%$ loss in this performance metric after training with about $9 \times
10^4$ samples. The best $\%$ performance loss obtained by the NNs is summarized
in Table \ref{table:cstrs}.

Second, we use the above trained NNs with the best performance losses in
validation simulations, with the setpoint and disturbance signals generated for
the unexpected plant behavior case. The $\%$ performance losses obtained in
these simulations are also summarized in the Table \ref{table:cstrs}, and the
performance degrades from less than $1\%$ to $5-8\%$. This deterioration in the
performance is due to some unseen ($x, u_{-1}, x_s, u_s$) encountered in
real-time by the NN resulting from the setpoint changes of the height of the
second CSTR, which was assumed to be fixed during the data generation. These
simulations illustrate that one must not expect the NNs to extrapolate outside
the training data. The practical implication of this study is that all possible
plant scenarios with the setpoints and disturbances must be identified, prior to
deployment of NNs, and should be used to generate the training data.

\begin{table*}[t]
    \caption{Summary of the simulation study for the industrial crude
    distillation unit example.}
    %\centering
\begin{tabular}{ |c|c|c|c|c|c|c|c| }
          \hline
          Controller & \thead{Neural network \\ architecture} & \thead{Number of
          \\ Parameters} & \thead{Memory \\ footprint \\ (MB)} & \thead{Training
          \\ time \\ (min)} & \thead{\% Performance\\ loss} & \thead{Average \\
          speedup} & \thead{Worst case \\ speedup} \\
          \hline
      SS &  &  & &  & 120.59 \% &  & \\ 
  satK &  &  &   &  & 13.07 \% & $4.83 \times 10^5$ & $1.51 \times 10^5$ \\ 
  SH &  &  &  &  & 1.56 \% & $6.66 \times 10^3$  & $2.51 \times 10^3$ \\ 
  NN-3-1664 & [1072, 1664, 1664, 1664, 32] & $1.85 \times 10^6$ & 14.18 & 52.66
  & 0.29 \% & $1.52 \times 10^4$ & $6.57 \times 10^3$  \\ 
  NN-3-1792 & [1072, 1792, 1792, 1792, 32]  & $2.11 \times 10^6$ & 16.15  &
  55.27 & 0.43 \% & $1.33 \times 10^4$ & $5.25 \times 10^3$ \\ 
  NN-3-1920 & [1072, 1920, 1920, 1920, 32]  & $2.39 \times 10^6$ & 18.24  &
  59.24 & 0.59 \% & $1.18 \times 10^4$ & $4.48 \times 10^3$ \\ 
  %NN-3-2048 & [1072, 2048, 2048, 2048, 32]  & $2.68 \times 10^6$ & 20.46  &
  %66.58 & 0.48 \% & $1.09 \times 10^4$ & $3.54 \times 10^3$\\ 
\hline
\end{tabular}
\label{table:cdu}      
\end{table*}

\begin{figure*}[!h]
    \includegraphics[page=41, width=0.5\textwidth,
        height=0.27\textheight]{cdu_comparision_plots.pdf} \hfill
        \includegraphics[page=43, width=0.5\textwidth,
        height=0.27\textheight]{cdu_comparision_plots.pdf} \vfill
    \vspace{-0.1in}
    \begin{center}\includegraphics[page=42, width=0.5\textwidth,
        height=0.27\textheight]{cdu_comparision_plots.pdf}\end{center}
    \vspace{-0.2in}
    \caption{Simulation statistics for the industrial crude distillation unit
    example. Data requirement to obtain accurate NN controllers (top left),
    online computation times of the NNs and MPC controller (top right), and the
    controller performance index obtained with the NNs, satK and MPC controllers
    (center bottom). The performance of the three NN and optimal MPC controllers
    are indistinguishable.}	
    \label{fig:cdu_statistics}
\end{figure*}

Third, the transient closed-loop performance of the trained NNs is examined for
one particular validation simulation for the expected plant behavior case.
Figures \ref{fig:cl_cstrs_inputs} and \ref{fig:cl_cstrs_outputs} show the
transient actuator and controlled measurements obtained with the NN-3-448
architecture, and optimal MPC controllers for a simulation period of 2 hours.
Figure \ref{fig:cstrs_statistics} (center bottom) shows the transient values of
the controller performance index obtained in this validation simulation with the
three structured NNs, SS, satK and optimal MPC controllers for the full
simulation period of 12 hours. Closed-loop performance obtained using NNs is
almost the same as the optimal MPC controller and the degradation in the
performance is quantified with the $\%$ performance loss metric.

Fourth, the online computational benefits of NNs over the QP based MPC is
analyzed. Figure \ref{fig:cstrs_statistics} (top right) shows a histogram of
online computation times of the NNs and QP solver in one validation simulation.
The average and worst-case speedups obtained using NNs are summarized in the
Table \ref{table:cstrs}. The QP solver takes about 8 to 13 seconds in this
example and all the NNs compute the control in just 0.1 to 4 milliseconds, and
are easily deployable in the sample time of 10 seconds.

The $\%$ performance loss obtained using the four heuristic controllers, their
corresponding speedups, maximum time required to train each NN, number of
parameters in each NN, and the memory required to store the weights of the NNs
is summarized in the Table \ref{table:cstrs}. The memory requirement shows that
NNs are deployable on resource constrained hardware, which can be crucial for
industrial deployment at regulatory control layers. The $85 \%$ performance loss
reported for the unstructured NN corresponds to training with $1.5 \times 10^5$
samples. Such performance is unacceptable for an industrial application. The
noticeable performance improvement with the structured NNs is because the
control law is exactly the same as the optimal feedback law at steady-states,
which simplifies the optimal feedback law approximation problem since the
training process only has to improve the NN control law for the transient
conditions.

\subsection{Large-scale crude distillation unit}

The scalability of the proposed NN design approach is next demonstrated on an
industrial crude distillation unit model \citep{pannocchia:rawlings:wright:2007}
with 252 states, 32 controls, and 90 measurements. The schematic of this model
is shown in Figure \ref{fig:schematic_cdu}, which is a typical crude unit found
in petrochemical refineries. The plant model is linear, and input and outputs
are scaled for the MPC regulator as in the previous example. Only four
measurements that represent the quality of the crude side products have
setpoints, and five disturbances are used on the crude composition, fuel gas
quality, and steam header pressure. The sample time for the measurements is 1
min. The tuning parameters and control horizon length for the MPC regulator are
chosen as $Q = 2C'C$, $R = 0.1I$, and $N = 140$. A total $3.6 \times 10^5$
samples of ($x, x_s, u_s, \kappa_N(\cdot)$) are generated for the NN training,
parallelized over 149 offline simulations. The time consumed for this data
generation step was 27.8 hours.

Three structured NN architectures are considered for this example. Each NN has
three hidden layers, with 1664 (NN-3-1664), 1792 (NN-3-1792), and 1920
(NN-3-1920) nodes in each hidden layer respectively. For the validation
simulations, one set of setpoint and disturbance signals for the expected plant
behavior case is generated with 24 setpoint changes and 48 disturbance changes
for 2880 timesteps (2 days). The short horizon MPC controller is constructed
with a horizon length of $N = 3$. The states and target steady-states ($x$,
$x_s$) are scaled for the NN training as in the previous example.

Figure \ref{fig:cdu_statistics} shows the statistics for the simulation study
performed for this example. The data requirement to obtain an accurate NN
controller is shown in the top left. The batch size in Adam was chosen as 2048
samples, and all the NNs are trained for 1500 epochs. The size of the buffer
dataset was chosen as 5$\%$ of the training dataset. The transient controller
performance index for one validation simulation obtained with the three
structured NNs, satK, and MPC controllers is shown in the center bottom. The top
right plot shows the histogram of online computation times of the NNs and QP
solver. Table \ref{table:cdu} gives a summary of the maximum time required to
train the NNs, number of parameters in each NN, memory footprint of the NNs, and
the $\%$ performance loss metric obtained by the structured NNs, SS, satK, and
SH controllers in the validation simulations.

The offline data generation and training times for the NNs shows that this
large-scale example is tractable with the proposed NN approach. This example is
challenging, since on average the QP solver CVXOPT requires 35 seconds, and 47
seconds in the worst-case. By comparison,  
all the NNs execute MPC in about 2 to 7 milliseconds. As observed from the plot
of the transient controller performance index, the performance degradation by
the NNs relative to the optimal MPC controller is negligible. The sampling of
the state space with the setpoint and disturbance signals for one particular
operational scenario, combined with the offset-free NN structure are the key
properties that enable scalability in this example.

\section{Conclusion} \label{sec:conclusion}

Neural networks have been demonstrated to approximate the MPC feedback law for
treating large-scale, linear MPC applications that may be out of reach with
online QP solvers. The proposed methods avoiding online optimization to date in
the literature, such as storing the entire MPC feedback law and sampling the
large dimensional state space for the NN training, are not tractable for the
problems considered in this article.

The key features for the scalability of the proposed approach are the structure
of the NN and data generation for only typical plant operational scenarios. The
NN design approach was demonstrated to scale on an industrial crude distillation
unit model with state and control dimensions of 252 and 32 respectively, and a
control-sample horizon length of 140. After the offline design phase, NNs
execute MPC about three to five orders of magnitude faster than an available QP
solver with less than 1$\%$ performance degradation in the closed-loop. The
deployment decision of the NNs must be based on the ability to reliably identify
and sample the anticipated plant operational scenarios with setpoints and large
magnitude disturbances that may change for a reasonable duration of operation.
Parallel computing can be used for the offline data generation step, if solving
QPs for each sampled state starts to take excessive amounts of time. After the
training process, the NNs are easily deployable on memory constrained hardware.
We established conditions under which the NN controllers are robust to state
estimation errors and process disturbances, providing some theoretical support
to practitioners interested in the deployment of NNs.

The online QP based MPC was adopted by practitioners due to its ability to
systematically handle multivariable systems and process constraints, and the
reliability of QP solution algorithms. The approach has guaranteed stability
properties. The need for sampling the plant operational scenarios and additional
computing hardware for training large applications are additional complexities
that must be considered in the design of NN controllers. The advantage of NNs is
their speed of MPC controller execution and convenience for deployment on memory
constrained hardware. Future research may be focused to establish conditions for
guaranteed closed-loop stability with NN controllers, systematically handling
state constraints, approximating the target selector QP, and extensions to
large, nonlinear MPC applications.

\section*{Acknowledgments}
The funding for this work was provided by the industrial members of the
Texas-Wisconsin-California Control Consortium (TWCCC). The computational
facilities purchased with funds from the National Science Foundation
(CNS-1725797, OAC-1925717) and administered by the Center for Scientific
Computing (CSC) was used for the simulation studies presented in this article.
The CSC is supported by the California NanoSystems Institute and the Materials
Research Science and Engineering Center (MRSEC; NSF DMR 1720256) at UC Santa
Barbara.

%% New version of the num-names style
\bibliographystyle{abbrvnat}
\bibliography{paper}

\section*{Appendix}
\renewcommand{\thesubsection}{\Alph{subsection}}

\subsection{Proof of Theorem \ref{thm:nnrobustness}} \label{app:theorem1}

We use the following definitions of robust positive invariance of a set with
respect to disturbances, input-to-state stability (ISS), ISS-Lyapunov function,
and the available result on the ISS property of a system if it admits an
ISS-Lyapunov function (\cite{jiang:wang:2001, rawlings:mayne:diehl:2017},
Appendix B).

\begin{definition} \label{def:robust_pos_invariance} (Robust positive
invariance). A closed set $X$ is robustly positive invariant for the system $x^+
= f(x, w)$, $w \in \bbW$ if $x \in X$ implies $f(x, w) \in X$ for all $w \in
\bbW$.
\end{definition}

\begin{definition} \label{def:iss} (Input-to-state stability (ISS)). $\bbW$ is a
compact set containing the origin and that $X$ is a closed robustly positive
invariant set for the system $x^+ = f(x, w)$, $w \in \bbW$. This system is ISS
in $X$ if there exist functions $\beta(\cdot) \in \mathcal{K}\mathcal{L}$ and
$\sigma(\cdot) \in \mathcal{K}$, such that for all $x \in X$ and $w(i) \in
\bbW$, $i \in \bbI_{0:k-1}$, $\norm{\psi(k; x, \mathbf{w}_k)} \leq
\beta(\norm{x}, k) + \sigma(\norm{\norm{\mathbf{w}_k}})$, in which $\psi(k; x,
\mathbf{w}_k)$ is the solution of the system at time $k$, starting from an
initial state $x$, and with the disturbance sequence $\mathbf{w}_k$ affecting
the system.
\end{definition}

\begin{definition} \label{def:iss_lyapunov_func} (ISS-Lyapunov function). A
function $V : X \rightarrow \bbR_{\geq 0}$ is an ISS-Lyapunov function in $X$
for the system $x^+ = f(x, w)$, $w \in \bbW$ if there exist functions
$\alpha_1(\cdot), \alpha_2(\cdot), \alpha_3(\cdot) \in \mathcal{K}_{\infty}$ and
$\sigma(\cdot) \in \mathcal{K}$ such that for all $x \in X$ and $w \in \bbW$, we
have
\begin{align*}
    \alpha_1(\norm{x}) \leq & V(x) \leq \alpha_2(\norm{x}), \\
    V(f(x, w)) &- V(x) \leq -\alpha_3(\norm{x}) + \sigma(\norm{w}).
\end{align*}
\end{definition}

\begin{prop} \label{prop:lyapunov_implies_iss} (ISS-Lyapunov function implies
ISS). $\bbW$ is a compact set containing the origin and that $X$ is a closed
robustly positive invariant set for $x^+ = f(x, w)$, $w \in \bbW$. If $f(\cdot)$
is continuous and there exists a continuous ISS-Lyapunov function in $X$ for the
system $x^+ = f(x, w)$, $w \in \bbW$, then the system is ISS in $X$.
\end{prop}

Additionally, we require the following two useful propositions from
\cite{allan:bates:risbeck:rawlings:2017} and \cite{rawlings:ji:2012}, to bound
values of continuous functions and an inequality for $\mathcal{K}$ functions
respectively.

\begin{prop} \label{prop:continuous_funcs} Let $C \subseteq D \subseteq \bbR^n$,
    $C$ compact, $D$ closed, and $f : D \rightarrow \bbR^n$ continuous.  
    Then there exists a $\sigma(\cdot) \in \mathcal{K}_{\infty}$, such that for
    all $x \in C$ and $y \in D$, we have that $\norm{f(x) - f(y)} \leq
    \sigma(\norm{x-y})$.
\end{prop}

\begin{prop} \label{prop:Kfunction_inequality} Let $\sigma(\cdot) \in
    \mathcal{K}$, the following holds for all $a_i \in \bbR_{\geq 0}$, $i \in
    \bbI_{1:n}$:
    \begin{align*}
        \sigma(a_1 + a_2 + \cdot \cdot + a_n) \leq \sigma(na_1) + 
        \sigma(na_2) + \cdot \cdot + \sigma(na_n)
    \end{align*}    
\end{prop}

\textbf{Proof of Theorem \ref{thm:nnrobustness}:} We proceed in two steps.
First, we show that there exist constants $\delta_1, \delta_2, \delta_3 > 0$
such that the set $S$ is robustly positive invariant. Then, we show that
$V_N^0(\xhat)$ is an ISS Lyapunov function for the perturbed system
\eqref{eq:perturbed_sys} on this set $S$. Denote $\tilde{x}^+ = A\xhat +
B\kappa_N(\xhat)$ as the nominal state evolution using the optimal MPC
controller, and $d(k) = [ e_{NN}(\xhat(k))'B', w(k)', e(k)', e(k+1)']'$ as the
combination of all the disturbances at a given time $k$.

To show robust positive invariance of $S$,  
we establish that if $V_N^0(\xhat) \leq \rho$, then $V_N^0(\xhat^+) \leq \rho$.
Since the optimal cost function ($V_N^0(\xhat)$) is continuous, using $\bbR^n$
as the closed set ($\xhat^+ \in \bbR^n$) and $S$ as the compact set ($\tilde{x}
\in S$), from Proposition \ref{prop:continuous_funcs} we have for some
$\sigma_1(\cdot) \in \mathcal{K}_{\infty}$ that
\begin{align}
    &\norm{V_N^0(\xhat^+) - V_N^0(\tilde{x}^+)} \leq 
    \sigma_1(\norm{\xhat^+ - \tilde{x}^+})  \nonumber  \\
    &\leq 
   \sigma_1(\norm{B e_{NN}(\xhat)} + \norm{w} + \norm{A}\norm{e} +  \norm{e^+}) \nonumber \\
     &\leq 
   \sigma_1(\norm{d} + \norm{d} + \norm{A}\norm{d} + \norm{d}) 
   := \sigma_2(\norm{d})
   \label{eq:optimal_cost_bound}
\end{align}
in which $\sigma_2(s) := \sigma_1(\norm{A} s +3s)$, and the last inequality
follows because $\norm{a} \leq \norm{[a', b']'}$, in which $a$ and $b$ are both
vectors, and $\sigma_2(\cdot) \in \mathcal{K}_{\infty}$. We partition $S$ in
outer and inner parts based on the value of the optimal cost $V_N^0(\xhat)$, and
bound the combination of disturbances $d$ in each part such that $V_N^0(\xhat^+)
\leq \rho$. \\
Case 1: $\rho/2 \leq V_N^0(\xhat) \leq \rho $. Using the upper bound on the
optimal cost, we have $\rho/(2c_2) \leq \norm{\xhat}^2$ for all $\xhat$ in this
part of $S$. To analyze the worst-case disturbance, we have from the equation
\eqref{eq:optimal_cost_bound} that $V_N^0(\xhat^+) \leq V_N^0(\tilde{x}^+) +
\sigma_2(\norm{d}) $. From nominal MPC cost decrease, we obtain  
$V_N^0(\xhat^+) \leq V_N^0(\xhat) - c_1\norm{\xhat}^2 + \sigma_2(\norm{d})$.
Thus, if $\sigma_2(\norm{d}) \leq \rho c_1/(2c_2)$, at all times in this outer
part of $S$, we have $V_N^0(\xhat^+) \leq \rho$. \\
Case 2: $V_N^0(\xhat) \leq \rho/2 $. From equation \eqref{eq:optimal_cost_bound}
and nominal MPC cost decrease, we have for this part of $S$ that $V_N^0(\xhat^+)
\leq \rho/2 + \sigma_2(\norm{d})$. Thus, if $\sigma_2(\norm{d}) \leq \rho/2$, at
all times in this inner part of $S$, we have $V_N^0(\xhat^+) \leq \rho$.

Therefore, if $\norm{d} \leq \textnormal{min} \{ \sigma_2^{-1}(\rho c_1/(2c_2)),
\sigma_2^{-1}(\rho/2)\} := \delta_{\textnormal{max}}$ at all times, then the set
$S$ is robustly positive invariant. Since $c_1 \leq c_2$,
$\delta_{\textnormal{max}} = \sigma_2^{-1}(\rho c_1/(2c_2))$. The constants
$\delta_1$ and $\delta_2$ are $\delta_{\textnormal{max}}/4$ each, and $\delta_3$
is $\delta_{\textnormal{max}}/(4\norm{B})$. The optimal MPC cost function
satisfies
\begin{align*}
    V_N^0(\xhat^+) - V_N^0(\xhat) &\leq 
    -c_1\norm{\xhat}^2 + \sigma_2(\norm{d})
\end{align*}
in the robustly positive invariant set $S$ and the disturbance set $\norm{d}
\leq \delta_{\textnormal{max}}$. Since the optimal cost function also satisfies
the upper and lower bounds $c_1\norm{x}^2 \leq V_N^0(x) \leq c_2\norm{x}^2$,
from the Definition \ref{def:iss_lyapunov_func}, it satisfies the requirements
of an ISS-Lyapunov function. Hence, using the Proposition
\ref{prop:lyapunov_implies_iss}, the perturbed linear system
\eqref{eq:perturbed_sys} is ISS (Definition \ref{def:iss}), and there exist
functions $\beta(\cdot) \in \mathcal{K}\mathcal{L}$ and $\sigma(\cdot) \in
\mathcal{K}$, such that $\norm{\phi_d(k; \xhat)} \leq \beta(\norm{\xhat}, k) +
\sigma(\norm{\norm{\mathbf{d}_k}})$. Using Proposition
\ref{prop:Kfunction_inequality} and the definition of $d(k)$, we have 
\begin{align*}
    \norm{\phi_d(k; \xhat)} & \leq \beta(\norm{\xhat}, k)
    + \sigma(\norm{B}\bar{e}_{NN} + \norm{\norm{\mathbf{w}_k}}
    + 2\norm{\norm{\mathbf{e}_{k+1}}}) \\
     & \leq \beta(\norm{\xhat}, k) \\
    & +\sigma(3\norm{B}\bar{e}_{NN}) + 
    \sigma(3\norm{\norm{\mathbf{w}_k}})
    + \sigma(6\norm{\norm{\mathbf{e}_{k+1}}})
\end{align*}
Therefore, the bound in the Theorem \ref{thm:nnrobustness} is established with
$\alpha_e(s) := \sigma(6s)$, $\alpha_w(s) := \sigma(3s)$, $\alpha_n(s) :=
\sigma(3\norm{B} s)$, and $\alpha_e(\cdot), \alpha_w(\cdot), \alpha_n(\cdot) \in
\mathcal{K}$.

\subsection{Parameters used in the CSTRs example} \label{app:cstrs_pars}

\renewcommand{\arraystretch}{1.1}
\begin{table}[!h]
    \centering
    \resizebox{\columnwidth}{!}{
    \begin{tabular}{|c|c|c|c|c|c|}
        \hline
        \multicolumn{6}{ |c| }{Parameters used to simulate the ODEs} \\
        \hline
        Parameter & Value & Unit & Parameter & Value & Unit\\
        \hline
        $\alpha_A$ & 3.5 &  & $k_m$ & 2.5 & \unit{m$^2$}\\ 
        $\alpha_B$ & 1.1 &  & $k_b$ & 1.5 & \unit{m$^2$} \\ 
        $\alpha_C$ & 0.5 &  & $\Delta H_1$ & -40 & \unitfrac{kJ}{kg} \\ 
        $\rho$ & 50 & \unitfrac{kg}{m$^3$} & $\Delta H_2$ & -50 &
        \unitfrac{kJ}{kg} \\ 
        $C_p$ & 3 & \unitfrac{kJ}{kg-K} & $E/R$ & 150 & \unit{K} \\ 
        $A_r$ & 0.3 & \unit{m$^2$} & $k^{\star}_1$ & $4 \times 10^{-4}$ &
        \unit{sec$^{-1}$} \\ 
        $A_m$ & 2 & \unit{m$^2$} & $k^{\star}_2$ & $1.8 \times 10^{-6}$ &
        \unit{sec$^{-1}$} \\ 
        $A_b$ & 4 & \unit{m$^2$} & $T_d$ & 313 & \unit{K} \\ 
        $k_r$ & 2.5 & \unit{m$^2$} &  &  &  \\ 
        \hline       
        \multicolumn{6}{ |c| }{Actuator constraints ($\underline{u},
        \overline{u}$)} \\
        \hline
        $F_0$ & (1.5, 2.5) & \unitfrac{kg}{sec} & $Q_r$ & (-500, 500) &
        \unit{kW} \\ 
        $F_1$ & (0.5, 1.5) & \unitfrac{kg}{sec} & $Q_m$ & (-500, 500) &
        \unit{kW} \\ 
        $D$ & (29.5, 30.5) & \unitfrac{kg}{sec} & $Q_b$ & (-500, 500) &
        \unit{kW} \\ 
        \hline
        \multicolumn{6}{ |c| }{Setpoint bounds ($\underline{r}_{sp},
        \overline{r}_{sp}$)} \\
        \hline
        $H_r$ & (158.8, 168.8) & \unit{m} & $T_r$ & (303, 323) & \unit{K} \\ 
        $H_m$ & (169.2, 179.2) & \unit{m} & $T_m$ & (310, 316) & \unit{K} \\ 
        $H_b$ & (2.2, 4.2) & \unit{m} & $T_b$ & (303, 323) & \unit{K} \\ 
        \hline
        \multicolumn{6}{ |c| }{Disturbance bounds ($\underline{d},
        \overline{d}$)} \\
        \hline
        $x_{A0}$ & (0.7, 0.85) & & $x_{B1}$ & (0, 0.15) & \\ 
        $x_{B0}$ & (0, 0.15) & & $T_0$ & (305, 321) & \unit{K} \\ 
        $x_{A1}$ & (0.7, 0.85) & & & & \\ 
        \hline       
        \multicolumn{6}{ |c| }{Steady state used for linearization ($x_s, u_s,
        d_s$)} \\
        \hline
        $H_r$ & 163.8 & \unit{m} & $F_0$ & 2 & \unitfrac{kg}{sec} \\ 
        $x_{Ar}$ & 0.40 &  & $F_1$ & 1 & \unitfrac{kg}{sec} \\ 
        $x_{Br}$ & 0.54 &  & $D$ & 30 & \unitfrac{kg}{sec} \\ 
        $T_r$ & 313.1 & \unit{K} & $Q_r$ & 0 & \unit{kW} \\ 
        $H_m$ & 174.2 & \unit{m} & $Q_m$ & 0 & \unit{KW} \\ 
        $x_{Am}$ & 0.37 &  & $Q_b$ & 0 & \unit{kW} \\ 
        $x_{Bm}$ & 0.58 & & $x_{A0}$ & 0.8 &  \\ 
        $T_m$ & 313.7 & \unit{K} & $x_{B0}$ & 0.1 & \\ 
        $H_b$ & 3.24 & \unit{m} &  $x_{A1}$&  0.8&  \\ 
        $x_{Ab}$ & 0.15 &  & $x_{B1}$ & 0.1 &  \\ 
        $x_{Bb}$ & 0.73 &  & $T_0$ & 313 & \unit{K} \\ 
        $T_b$ & 313.7 & \unit{K} & & & \\ 
        \hline
    \end{tabular}} \caption{Parameters used in the ODEs to simulate the plant in
    the CSTRs in series with a flash separator example, actuator constraints,
    setpoint and disturbance bounds, and the steady state used to obtain the
    linear model for the MPC controller.}
\label{table:cstrs_pars}
\end{table}

\end{document}